### 2004-07-05 Documentation/part1.tex
*** empty log message ***

\begin{table}[t]
\begin{center}
\begin{tabular}{p{3.8cm}p{3.8cm}p{7cm}}
\hline
Function & Purpose & Parameters \ra\ Result \\
\hline
\multicolumn{3}{l}{{\bf Systematics only:}} \\
\GLB{glbChiSys} & $\chi^2$ with systematics only  & ({\tt glb\_params in, int exp, int rule}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
\multicolumn{3}{l}{{\bf Projections onto axes:}} \\
\GLB{glbChiTheta} & Projection onto $\theta_{13}$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
\GLB{glbChiDelta} & Projection onto $\deltacp$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
\GLB{glbChiTheta23} & Projection onto $\theta_{23}$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
\GLB{glbChiDm} & Projection onto $\ldm$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
\GLB{glbChiDms} & Projection onto $\sdm$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
\multicolumn{3}{l}{{\bf Projection onto plane:}} \\
\GLB{glbChiThetaDelta} & Projection onto $\theta_{13}$-$\deltacp$-plane  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
\multicolumn{3}{l}{{\bf Projection onto any hyper-plane:}} \\
\GLB{glbChiNP} & Projection onto any $n$-dimensional hyper-plane  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \newline
Needs \GLB{glbSetProjection} before! \\[0.2cm]
\multicolumn{3}{l}{{\bf Localization of degeneracies:}} \\
\GLB{glbChiAll} & (Local) Minimization over all parameters  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\
\hline
\end{tabular}
\end{center}
\caption{\label{tab:stdfunctions} \index{Standard functions (table)} The \GLOBES\ standard function to obtain a $\chi^2$-value with systematics only or systematics and correlations. The parameters {\tt rule} and {\tt exp}
can either be \GLB{GLB\_ALL} for all initialized experiment or the
experiment number ($0$ to \GLB{glb\_num\_of\_exps}-1) for a specific experiment. The format of \GLB{glb\_params} is discussed in detail in \Chapt~\ref{chapt:gettingstarted}. Note that all functions but {\tt ChiSys}
  are using minimizers which have to be initialized with \GLB{glbSetInputErrors} and \GLB{glbSetStartingValues} first.}
\end{table}

@@

Similarly, the density parameters or iterations can be accessed:
\begin{function}
\GLBNS{glbSetDensityParams}
{\tt glb\_params glbSetDensityParams(glb\_params in, double dens, int which)} sets the density parameter {\tt which} in the structure {\tt in}
to the value {\tt dens}. If the assignment was unsuccessful, the 
function returns {\tt NULL}.
\end{function}
\begin{function}
\GLBNS{glbGetDensityParams}
{\tt double glbGetDensityParams(glb\_params in, int which)}
returns the value of the density parameter {\tt which} in the 
structure {\tt in}.
\end{function}
\begin{function}
\GLBNS{glbSetIteration}
{\tt glb\_params glbSetIteration(glb\_params in, int iter)} sets the number of iterations in the structure {\tt in}
to the value {\tt iter}. If the assignment was unsuccessful, the 
function returns {\tt NULL}.
\end{function}
\begin{function}
\GLBNS{glbGetIteration}
{\tt int glbGetIteration(glb\_params in)}
returns the value of the number of iterations in the 
structure {\tt in}.
\end{function}
In total, the parameter vector handling in a program normally has the
following order:
\begin{quote}
{\tt
glbInitExperiment(...); \\
/* ... more initializations ... */ \\
\\
glb\_params vector1 = glbAllocParams(); \\
/* ... more vectors allocated ... */ \\
\\
/* Program code: assign and use vectors */ \\
\\
glbFreeParams(vector1); \\
/* ... more vectors destroyed ... */ \\
\\
/* ... end of program or glbClearExperimentList ... */ 
}
\end{quote}

@@

\index{External input}
It is one of the strengths of the \GLOBES\ software to use external input to reduce the fit manifold with the knowledge from external (eariler) measurements. The treatment of external input is done by the addition of Gaussian {\em priors} to the final\footnote{After the systematics minimization and after all $\chi^2$'s are added for all channels} $\chi^2$-function  and the minimization over the respective parameters. For example, for the matter density, one obtains as the minimized $\chi^2_F$ after minimzation over $\rho$
\be
 \chi^2_F = \underset{\rho}{\mathrm{min}} \left( \chi^2(\rho) + \frac{(\rho - \rho^0)^2}{\sigma_\rho^2} \right).
\label{equ:priors}
\ee
In practice, this minimzation is done simultaneously over all priors and free oscillation parameters, but this example serves as a very simple illustration. In \equ{priors}, $\rho^0$ is the {\em starting value} of the prior, and $\sigma_\rho$ the $1 \sigma$ absolute {\em input error}. Thus, it is assumed that an external measurement has determined the matter density with a precision (input error) $\sigma_\rho$ at the central value $\rho^0$. Usually, the starting value corresponds to the best-fit value and the input error to the $1 \sigma$ half width of the external measurement. For the matter density, $\rho^0$ can be set to the average matter density $\bar{\rho}$ and $\sigma_\rho$ to the matter density uncertainty. Since \GLOBES\ may also use a matter density profile, it actually uses a scaling factor $\hat\rho$ instead of the average matter density (see last section), \ie, $\rho^0=1.0$ is the central value in either case. Thus, $\sigma_\rho$ directly correponds to the uncertainty of this scaling factor. For instance, using the PREM (``Preliminary Reference Earth Model'') profile, $\rho^0 = 1.0$ and $\sigma_\rho = 0.05$ is a conservative estimate of the PREM profile uncertainty.  \index{PREM profile}

In principle, one can set the priors for the matter density and all oscillation parameters. For example, if the disappearance channels of the experiment determine the leading oscillation parameters, once can set very large values for the input errors ($\sigma$'s). If, however, earlier external measurements provide better information, one can set their absolute precisions with the input errors. The starting values are usually set to the best-fit values. In some cases, it may be necessary to adjust them, such as for $\ldm$ and the negative mass hierarchy to $\rho^0_{\ldm} = - |\ldm|$ if some external precision on $| \ldm |$ is imposed and one is investigating the negative-sign solution. In other cases, minor modifications of the starting values can cause a faster convergence of the algorithm.
In either case, two function have to be called {\em before the usage of any minimizer}:
\begin{function}
\GLB{SetStartingValues}$(\{\rho^0_{\theta_{12}}, \, \rho^0_{\theta_{13}}, \, \rho^0_{\theta_{23}}, \, \rho^0_{\deltacp}, \, \rho^0_{\sdm}, \, \rho^0_{\ldm}, \, \rho^0_{\hat\rho} \})$ sets the starting values for all of the following minimizer calls.
\end{function}
\begin{function}
\GLB{SetInputErrors}$(\{\sigma_{\theta_{12}}, \, \sigma_{\theta_{13}}, \, \sigma_{\theta_{23}}, \, \sigma_{\deltacp}, \, \sigma_{\sdm}, \, \sigma_{\ldm}, \, \sigma_{\hat\rho} \})$ sets the input errors for all of the following minimizer calls.
\end{function}
Both functions may take one matter density parameter or as many as there are experiments. Since the assumptions about the matter density profile are usually similar for the loaded experiments, using only one matter density parameter is a useful tool to apply the same starting value  and input error to all loaded experiments.
Eventually, a typical initialization of the external input may look like this:
\begin{quote}
{\tt
SetStartingValues(\{th12, th13, th23, dcp, sdm, ldm, 1.0\}); \\
SetInputErrors(\{th12*0.1, 10, 10, 10, sdm*0.1, ldm, 0.05\}); 
}
\end{quote}
In this example, the starting values are set to the best-fit values. The input errors for $\theta_{13}$, $\theta_{23}$, $\deltacp$, and $\ldm$ are kept large, since the experiment measures these parameters itself. However, some precision on $\ldm$ is imposed to avoid ending up in the negative-sign solution. The solar parameters $\theta_{12}$ and $\sdm$ are assumed to be known with $10\%$ precision each.\footnote{In fact, accelerator-based long-baseline experiments are only sensitive to the product $\sin 2 \theta_{12} \cdot \sdm$, which means that these errors effectively add up to an error of this product.} Finally, the matter density uncertainties are assumed to have an amplitude of $5\% \,  \bar\rho$.

@@

\begin{quote}
{\tt {\scriptsize
  /* Set starting values and input errors for all projections */  \\
  glbDefineParams(input\_errors,theta12*0.1,10,10,10,sdm*0.1,ldm);\\  
  glbSetDensityParams(input\_errors,0.05,GLB\_ALL);\\
  glbSetStartingValues(true\_values);\\
  glbSetInputErrors(input\_errors);\\
\\
  /* Define two-parameter projection: Only deltacp is free! */ \\
  glbDefineProjection(th13\_projection,GLB\_FIXED,GLB\_FIXED,GLB\_FIXED,GLB\_FREE,GLB\_FIXED,GLB\_FIXED); \\
  glbSetProjection(th13\_projection); \\ 
\\
  /* Iteration over all values to be computed */ \\
  double x,res1,res2;     \\
  for(x=-4;x<-2.0+0.001;x=x+2.0/50) \\
  \{ \\
\hspace*{0.5cm}  /* Set fit value of stheta */ \\
\hspace*{0.5cm} glbSetOscParams(test\_values,asin(sqrt(pow(10,x)))/2,1); \\
    \\
\hspace*{0.5cm} /* Guess fit value for deltacp in order to safely find minimum */ \\
\hspace*{0.5cm} glbSetOscParams(test\_values,200.0/2*(x+4)*M\_PI/180,3); \\
 \\
\hspace*{0.5cm} /* Compute Chi2 for user-defined two-parameter correlation */ \\
\hspace*{0.5cm} res1=glbChiNP(test\_values,NULL,GLB\_ALL); \\
      \\
\hspace*{0.5cm} /* Compute Chi2 for full correlation: minimize over all but theta13 */ \\
\hspace*{0.5cm} res2=glbChiTheta(test\_values,NULL,GLB\_ALL); \\
  \\
\hspace*{0.5cm} AddToOutput(x,res1,res2);\\
  \} \\

@@

The projection onto the $\stheta$- (or $\deltacp$-) axis is performed by fixing $\stheta$ (or $\deltacp$) and minimizing the $\chi^2$-function over all free fit parameters and the matter densities. We illustrate this method at the example of the projection of the two-dimensional manifold in the $\stheta$-$\deltacp$-plane onto the $\stheta$-axis in \figu{projex}. In this figure, the left-hand plot shows the correlation in the $\stheta$-$\deltacp$-plane computed with {\tt Chi} or {\tt SingleChi}. The right-hand plot shows the projection of this two-dimensional manifold onto the $\stheta$ axis by minimizing $\chi^2$ over $\deltacp$. In this simple example, the minimization is done along the vertical gray lines in the left hand plot. The obtained minima are located on the thick gray curve, which means the the right-hand plot represents the $\chi^2$-value along this curve. In fact, one can easily see that one obtains the correct projected $3 \sigma$ errors in this example (\cf, arrows). This figure illustrates the projection of a two-parameter correlation. In general, the full $n$-parameter correlation is treated similarly by the simultaneous (local) minimization over all free fit parameters. The example on page~\pageref{ex:corrproj} demonstrates how one can obtain \figu{projex} (right) with keeping all parameters but $\deltacp$ fixed, as well as how one can inlcude the full $n$-parameter correlation with external input. It also demonstrates how these two compare to each other.

One mentioned disadvantage of the local minimizer is the possibility to end up in a local minimum. Usually it works very well to start the minimizer close to the best-fit values. In the example in \figu{projex}, the minimizer is always started at $\deltacp=90^\circ$ and finds its way to the appropriate minimum. In some cases, however, the topology may be more difficult and the true minimum is not correctly found. A strong indication for such a situation are discontinuities in the projected $\chi^2$-function, where the minimizer jumps from one minimum to the other. In such a case, the starting point of the minimizer has to be adjusted to help it find the true minimum. In other cases, the convergence of the minimizer may be infinitively small. Then it usually helps to start the minimizer somewhat off the best-fit values. The following functions are the simplest minimizers provided by \GLOBES :
\begin{function}
\GLB{ChiTheta}$( \theta_{13}, \, \{ \theta_{12}, \, \theta_{23}, \, \deltacp , \, \sdm , \, \ldm, \, \hat{\rho_1}, \,  \hdots ,  \,\hat{\rho_n} \})$ returns the projection onto the $\theta_{13}$-axis for all experiments. The parameters are the fixed value of $\theta_{13}$ and the starting point of the minimizer including all starting values for the matter density scaling factors (without $\theta_{13}$, which is fixed). The return value is a list $\{ \chi^2, \, \theta_{12}, \, \theta_{23}, \, \deltacp , \, \sdm , \, \ldm, \, \hat{\rho_1}, \, \hdots , \, \hat{\rho_n} , \, N_{\mathrm{Iter}} \}$ with the minimum $\chi^2$ found, the coordinates of the local minimum (without $\theta_{13}$), and the number of iterations $N_{\mathrm{Iter}}$ used by the minimizer (number of function calls of {\tt Chi}).
\end{function}
\begin{function}
\GLB{SingleChiTheta}$( \theta_{13}, \, \{ \theta_{12}, \, \theta_{23}, \, \deltacp , \, \sdm , \, \ldm,  \, \hat{\rho}_{N_{\mathrm{exp}}} \}, \, N_{\mathrm{exp}})$ works similar to {\tt ChiTheta}, but only for one of the loaded experiments $N_{\mathrm{exp}}$. It returns the list $\{ \chi^2, \, \theta_{12}, \,  \theta_{23}, \, \deltacp , \, \sdm , \, \ldm, \, \hat{\rho}_{N_{\mathrm{exp}}} , \, N_{\mathrm{Iter}} \}$.
\end{function}
\begin{function}
\GLB{ChiDelta}$( \deltacp, \, \{ \theta_{12}, \, \theta_{13}, \, \theta_{23}, \, \sdm , \, \ldm, \, \hat{\rho_1}, \,  \hdots ,  \,\hat{\rho_n} \})$ returns the projection onto the $\deltacp$-axis for all experiments. The parameters are the fixed value of $\deltacp$ and the starting point of the minimizer including all starting values for the matter density scaling factors (without $\deltacp$, which is fixed). The return value is a list $\{ \chi^2, \, \theta_{12}, \, \theta_{13}, \, \theta_{23}, \, \sdm , \, \ldm, \, \hat{\rho_1}, \, \hdots , \, \hat{\rho_n} , \, N_{\mathrm{Iter}} \}$ with the minimum $\chi^2$ found, the coordinates of the local minimum (without $\deltacp$), and the number of iterations $N_{\mathrm{Iter}}$ used by the minimizer (number of function calls of {\tt Chi}).
\end{function}
\begin{function}
\GLB{SingleChiDelta}$( \deltacp, \, \{ \theta_{12}, \, \theta_{13}, \, \theta_{23}, \, \sdm , \, \ldm, \,  \hat{\rho}_{N_{\mathrm{exp}}} \}, \, N_{\mathrm{exp}})$ works similar to {\tt ChiDelta}, but only for one of the loaded experiments $N_{\mathrm{exp}}$. It returns the list $\{ \chi^2, \, \theta_{12}, \,  \theta_{13}, \,  \theta_{23}, \, \sdm , \, \ldm, \, \hat{\rho}_{N_{\mathrm{exp}}} , \, N_{\mathrm{Iter}} \}$.
\end{function}
All of these function have the same parameter structure: The fixed parameter is not transferred in the list, but as a the first separate parameter. The number of matter density scaling factors corresponds to the number of experiments used, since each experiment may face other matter density conditions. Note that before any of these function calls, {\tt SetStartingValues} and {\tt SetInputErrors} have to be used at least once. In addition, note that the resulting $\chi^2$ of {\tt ChiTheta} (or {\tt ChiDelta}) is not the sum of the $\chi^2$-values over all {\tt Single}-functions of all experiments anymore. This has two reasons: First, the topology of the fit manifold is altered by the addition of $\chi^2$-values of different experiments. Thus, after the minimization, the position of the minimum can be different to the ones of the individual experiments. Second, the priors for the external knowledge on the parameters are only added once -- independent of the number of experiments. A simple application of {\tt ChiTheta} can be found in the example on page~\pageref{ex:corrproj}.

>>>>

\begin{table}[tpb]
\begin{center}
\begin{tabular}{p{3.8cm}p{3.8cm}p{7cm}}
\hline
Function & Purpose & Parameters \ra\ Result \\
\hline
\multicolumn{3}{l}{{\bf Systematics only:}} \\
\GLB{glbChiSys} & $\chi^2$ with systematics only  & ({\tt glb\_params in, int exp, int rule}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
\multicolumn{3}{l}{{\bf Projections onto axes:}} \\
\GLB{glbChiTheta} & Projection onto $\theta_{13}$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
\GLB{glbChiDelta} & Projection onto $\deltacp$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
\GLB{glbChiTheta23} & Projection onto $\theta_{23}$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
\GLB{glbChiDm} & Projection onto $\ldm$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
\GLB{glbChiDms} & Projection onto $\sdm$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
\multicolumn{3}{l}{{\bf Projection onto plane:}} \\
\GLB{glbChiThetaDelta} & Projection onto $\theta_{13}$-$\deltacp$-plane  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
\multicolumn{3}{l}{{\bf Projection onto any hyper-plane:}} \\
\GLB{glbChiNP} & Projection onto any $n$-dimensional hyper-plane  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \newline
Needs \GLB{glbSetProjection} before! \\[0.2cm]
\multicolumn{3}{l}{{\bf Localization of degeneracies:}} \\
\GLB{glbChiAll} & (Local) Minimization over all parameters  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\
\hline
\end{tabular}
\end{center}
\caption{\label{tab:stdfunctions} \index{Standard functions (table)} The \GLOBES\ standard function to obtain a $\chi^2$-value with systematics only or systematics and correlations. The parameters {\tt rule} and {\tt exp}
can either be \GLB{GLB\_ALL} for all initialized experiment or the
experiment number ($0$ to \GLB{glb\_num\_of\_exps}-1) for a specific experiment. The format of \GLB{glb\_params} is discussed in detail in \Chapt~\ref{chapt:gettingstarted}. Note that all functions but {\tt ChiSys}
  are using minimizers which have to be initialized with \GLB{glbSetInputErrors} and \GLB{glbSetStartingValues} first.}
\end{table}

@@

Similarly, the density parameters or iterations can be accessed:
\begin{function}
\GLBNS{glbSetDensityParams}
{\tt glb\_params glbSetDensityParams(glb\_params in, double dens, int which)} sets the density parameter {\tt which} in the structure {\tt in}
to the value {\tt dens}. If the assignment was unsuccessful, the 
function returns {\tt NULL}. If {\tt GLB\_ALL} is used for {\tt which},
the density parameters of all experiments will be set accordingly.
\end{function}
\begin{function}
\GLBNS{glbGetDensityParams}
{\tt double glbGetDensityParams(glb\_params in, int which)}
returns the value of the density parameter {\tt which} in the 
structure {\tt in}.
\end{function}
\begin{function}
\GLBNS{glbSetIteration}
{\tt glb\_params glbSetIteration(glb\_params in, int iter)} sets the number of iterations in the structure {\tt in}
to the value {\tt iter}. If the assignment was unsuccessful, the 
function returns {\tt NULL}.
\end{function}
\begin{function}
\GLBNS{glbGetIteration}
{\tt int glbGetIteration(glb\_params in)}
returns the value of the number of iterations in the 
structure {\tt in}.
\end{function}
In total, the parameter vector handling in a program normally has the
following order:
\begin{quote}
{\tt
glbInitExperiment(...); \\
/* ... more initializations ... */ \\
\\
glb\_params vector1 = glbAllocParams(); \\
/* ... more vectors allocated ... */ \\
\\
/* Program code: assign and use vectors */ \\
\\
glbFreeParams(vector1); \\
/* ... more vectors destroyed ... */ \\
\\
/* ... end of program or glbClearExperimentList ... */ 
}
\end{quote}

@@

\index{External input}
It is one of the strengths of the \GLOBES\ software to use external input in order to reduce the fit manifold with the knowledge from external
(eariler) measurements. The treatment of external input is done by the addition of Gaussian so-called {\em priors} to the systematics-minimized $\chi^2$-function. For example, for the matter density, one obtains as the final projected $\chi^2_F$ after minimzation over the matter density
scaling factor \index{Matter density scaling factor} $\hat{\rho}$
\be
 \chi^2_F = \underset{\rho}{\mathrm{min}} \left( \chi^2(\hat{\rho}) + \frac{(\hat{\rho} - \hat{\rho}^0)^2}{\sigma_{\hat{\rho}}^2} \right).
\label{equ:priors}
\ee
This example is a very simple one, since in fact the
minimzation is simultaneously performed over all priors and free oscillation parameters. In \equ{priors}, $\hat{\rho}^0$ is the {\em starting value} of the prior, and $\sigma_{\hat{\rho}}$ the $1 \sigma$ absolute (half width) {\em input error}. Thus, it is assumed that an external measurement has determined the matter density with a precision (input error) $\sigma_{\hat{\rho}}$ at the central value $\hat{\rho}^0$. Usually, the starting value is fixed at the best-fit value, and the input error to the $1 \sigma$ half width of the external measurement. For the matter density, $\hat{\rho}^0$  is usually set to $1.0$ corresponding to
the actual matter density profile such as given by the experiment
 definition file, and $\sigma_{\hat{\rho}}$ to the 
relative matter density uncertainty (\eg, $0.05$ for 5\% uncertainty).

In principle, one can set the priors for the matter density and all oscillation parameters. For example, if the disappearance channels of the experiment determine the leading oscillation parameters with unprecedented
precisions, one can omit the respective input errors. In \GLOBES , a
value of $0$ correponds to neglecting the prior. If, however, earlier external measurements provide better information, one can set their absolute precisions with the input errors. The starting values are usually chosen to be the best-fit values of this external input, such as
for the precisions of the solar experiments. In some cases, it may be
necessary to adjust them, such as for artificial contraints to the
oscillation parameters. In other cases, minor modifications of the starting values can cause a faster convergence of the algorithm.
For example, for the investigation of the
opposite-sign solution, one can use the prior to constrain $\ldm$
in order to force the minimizer not to  fall into the (unwanted) best-fit solution. In this case, the starting value of $\ldm$ would be set to $\rho^0_{\ldm} = - \ldm$, and a $\sigma_{\ldm}$ of the order of $\ldm$ would be imposed. For the algorithm, it would then be rather difficult 
to converge into the unwanted best-fit solution. However, note that one should in this case check that the actually determined value for $\ldm$ after minimization is close enough to the guessed values $-\ldm$ in order to avoid significant artifical contributions of the priors to the 
final $\chi^2$.

In order to set the starting values and input errors, two function have to
 be called {\em before the usage of any minimizer}:
\begin{function}
\GLBNS{glbSetStartingValues}
{\tt int glbSetStartingValues(const glb\_params in)} sets the starting values for all of the following minimizer calls to {\tt in}.
\end{function}
\begin{function}
\GLBNS{glbSetInputErrors}
{\tt int glbSetInputErrors(const glb\_params in)}
 sets the input errors for all of the following minimizer calls to {\tt in}.
 An input error of $0$ correponds to not taking into account the
 respective prior.
\end{function}
Accordingly, there are functions to return the actually set starting values
and input errors:
\begin{function}
\GLBNS{glbGetStartingValues}
{\tt int glbGetStartingValues(glb\_params out)} writes the currently
set starting values to {\tt out}.
\end{function}
\begin{function}
\GLBNS{glbGetInputErrors}
{\tt int glbGetInputErrors(glb\_params out)}
 writes the currently set input errors to {\tt out}.
\end{function}
All functions take or return as many matter density parameters as there are initialized experiments. In addition, they return $-1$ if the operation
was not successful.

Eventually, a typical initialization of the external input with
$10\%$ external precisions for the solar parameters\footnote{In fact, accelerator-based long-baseline experiments are primarily sensitive to the product $\sin 2 \theta_{12} \cdot \sdm$, which means that these errors effectively add up to an error of this product of about $15\%$.}, 
and $5\%$ matter density uncertainties for all experiments looks like this:
\begin{quote}
{\tt
 glbDefineParams(input\_errors,theta12*0.1,0,0,0,sdm*0.1,0);\\  
 glbSetDensityParams(input\_errors,0.05,GLB\_ALL);\\
 glbSetStartingValues(true\_values);\\
 glbSetInputErrors(input\_errors);\\
}
\end{quote}
In this example, the starting values are set to the best-fit (simulated) values. Remember that initially the matter density scaling factors are 
all $1.0$, which means that they do not need to be adjusted for the
starting values.

Though the priors are an elegant way to treat external input, there are also some complications with priors. The following hints are for the
more advanced \GLOBES\ user:
\begin{enumerate}
\item
 The priors are only added once to the final $\chi^2$, no matter how
 many experiments there are simulated. This is already one reason
 (besides the minimization) why the sum of all projected $\chi^2$'s of 
 the indivual experiments 
 cannot correspond to the $\chi^2$ of the combination of all experiments.
\item
 Priors are not used for parameters which are not minimized over, \ie, 
 kept fixed. This will be important together with arbitrary projections
 using \GLB{glbChiNP}. A more subtile consequence is the comparison
 of fit manifold sections and projections for the solutions where
 the absolute minimum $\chi^2$ is larger than zero, \ie, degeneracies
 other than the best-fit solution. In this case, the sections and
 projections are not  comparable if not corrected by the prior contributions, where the
 correction can be obtained as the $\chi^2$-difference at the minimum.
 For example, projecting the $\mathrm{sgn}(\ldm)$-degeneracy onto
 the $\theta_{13}$-$\deltacp$-plane and comparing it with the section
 (all other parameters fixed), the section region would in many cases be 
 larger than the projection region if the priors are not added to the
 section. At the best-fit solution, this problem usually does not occur
 because the prior contributions are close to zero.    
\item
Currently, \GLOBES\ only supports Gaussian priors for the individual
oscillation parameters. Especially for the solar parameters, this is only
 an approximation, since they are imposed on $\theta_{12}$ and not
 on $\sin 2 \theta_{12}$, $\sin 2 \theta_{12} \cdot \sdm$, or $\sin  
 \theta_{12}$. Later versions of  \GLOBES\ may include more alternatives.
 \end{enumerate}

@@

\begin{quote}
{\tt {\scriptsize
  /* Set starting values and input errors for all projections */  \\
  glbDefineParams(input\_errors,theta12*0.1,0,0,0,sdm*0.1,0);\\  
  glbSetDensityParams(input\_errors,0.05,GLB\_ALL);\\
  glbSetStartingValues(true\_values);\\
  glbSetInputErrors(input\_errors);\\
\\
  /* Define two-parameter projection: Only deltacp is free! */ \\
  glbDefineProjection(th13\_projection,GLB\_FIXED,GLB\_FIXED,GLB\_FIXED,GLB\_FREE,GLB\_FIXED,GLB\_FIXED); \\
  glbSetProjection(th13\_projection); \\ 
\\
  /* Iteration over all values to be computed */ \\
  double x,res1,res2;     \\
  for(x=-4;x<-2.0+0.001;x=x+2.0/50) \\
  \{ \\
\hspace*{0.5cm}  /* Set fit value of stheta */ \\
\hspace*{0.5cm} glbSetOscParams(test\_values,asin(sqrt(pow(10,x)))/2,1); \\
    \\
\hspace*{0.5cm} /* Guess fit value for deltacp in order to safely find minimum */ \\
\hspace*{0.5cm} glbSetOscParams(test\_values,200.0/2*(x+4)*M\_PI/180,3); \\
 \\
\hspace*{0.5cm} /* Compute Chi2 for user-defined two-parameter correlation */ \\
\hspace*{0.5cm} res1=glbChiNP(test\_values,NULL,GLB\_ALL); \\
      \\
\hspace*{0.5cm} /* Compute Chi2 for full correlation: minimize over all but theta13 */ \\
\hspace*{0.5cm} res2=glbChiTheta(test\_values,NULL,GLB\_ALL); \\
  \\
\hspace*{0.5cm} AddToOutput(x,res1,res2);\\
  \} \\

@@

The projection onto the $\stheta$- (or $\deltacp$-) axis is performed by fixing $\stheta$ (or $\deltacp$) and minimizing the $\chi^2$-function over all free fit parameters and the matter densities. We illustrate this method at the example of the projection of the two-dimensional manifold in the $\stheta$-$\deltacp$-plane onto the $\stheta$-axis in \figu{projex}. In this figure, the left-hand plot shows the correlation in the $\stheta$-$\deltacp$-plane computed with {\tt glbChiSys}. The right-hand plot illustrates the projection of this two-dimensional manifold onto the $\stheta$ axis by minimizing $\chi^2$ over $\deltacp$. In this simple example, the minimization is done along the vertical gray lines in the left hand plot. The obtained minima are located on the thick gray curve, which means the the right-hand plot represents the $\chi^2$-value along this curve. In fact, one can easily see that one obtains the correct projected $3 \sigma$ errors in this example (\cf, arrows). This figure illustrates the projection of a two-parameter correlation. In general, the full $n$-parameter correlation is treated similarly by the simultaneous (local) minimization over all free fit parameters.

The following functions are some of the simplest minimizers 
provided by \GLOBES :
\begin{function}
\GLBNS{glbChiTheta}
{\tt double glbChiTheta(const glb\_params in, glb\_params out, int exp)}
returns the projected $\chi^2$ onto the $\theta_{13}$-axis for the  experiment {\tt exp}. For the simulation of all initialized experiments,
use {\tt GLB\_ALL} for {\tt exp}. The values in {\tt in} are the guessed fit values for the minimizer (all parameters other than $\theta_{13}$) and the fixed fit value of $\theta_{13}$. The actually determined parameters at the minimum are returned in {\tt out}, where $\theta_{13}$ is still at
its fixed value. If {\tt out} is set to {\tt NULL}, this information will not be returned.
\end{function}
\begin{function}
\GLBNS{glbChiDelta}
{\tt double glbChiDelta(const glb\_params in, glb\_params out, int exp)}
returns the projected $\chi^2$ onto the $\deltacp$-axis for the  experiment {\tt exp}. For the simulation of all initialized experiments,
use {\tt GLB\_ALL} for {\tt exp}. The values in {\tt in} are the guessed fit values for the minimizer (all parameters other than $\deltacp$) and the fixed fit value of $\deltacp$. The actually determined parameters at the minimum are returned in {\tt out}, where $\deltacp$ is still at its fixed value. If {\tt out} is set to {\tt NULL}, this information will not be returned.
\end{function}

All of the minimization functions have a similar parameter structure: The fixed fit parameter value and the guessed starting point of the minimizer, \ie, the guessed position of the minimum, are transferred in the list {\tt in}. Part of this list are the matter density
scaling factors of all experiments, which are also minimized over. 
The minimizer is then started at the guessed point and runs into
the local minimum, where the fit parameter of the projection
axis is fixed. For the best-fit solution, it is usually sufficient to start the minimizer at the best-fit point. However, the convergence speed might
be better by starting it slightly off this point. In addition, there
are problems in many cases with more complicated topologies, which means
that better guesses for the position of the minimum are needed.
The position of the minimum is then returned in {\tt out}
together with the number of iterations used for the minimization.
It is very often useful to print the output of the minimization with
\GLB{glbPrintParams} in order to check that the minimum is the
appropriate one. For example, if the mininizer ends up in the wong-sign
solution in $\ldm$, priors can be used to force it into the tested
minimum. In addition, the number of iterations used allows an optimization
of the convergence speed. 
% 
Note that before any minimization, \GLB{glbSetStartingValues} 
and \GLB{glbSetInputErrors} have to be used at least once. In addition, note that the resulting $\chi^2$ of {\tt glbChiTheta} (or {\tt glbChiDelta}) for the combination of more than one experiment is not 
equal to the sum of the individual $\chi^2$-values anymore. This has two reasons: First, the topology of the fit manifold is altered by the addition of $\chi^2$-values of different experiments. Thus, after the minimization, the position of the minimum can be different to the ones of the individual experiments. Second, the priors for the external knowledge on the parameters are only added once -- independent of the number of experiments.

\index{Matter density scaling factor}
The the output of the minimizer in {\tt out} carries as many matter density scaling factors as there are experiments. Either
one (for the simulation of one experiment) or all (for the simulation of
all experiments) are different from $1.0$ if matter density uncertainties
are present, since each experiment may face other matter density conditions. The minimizer for indivdual experiments ``know'' which
experiment they are currently treating, which means that only the returned
matter density scaling factor of the appropriate experiment is modified.
For example, calculating {\tt glbChiTheta} for the last experiment
number, the last density values will be modified.
This approach turns out to be extremely useful together with the
simulation of more than one experiment. One can, for instance, locate the
degeneracies of all individual experiments. In order to test if
these degeneracies are still present in the combination of all experiments (which has a very different topology), one can test the combination of experiments with the output {\tt out} from the individual experiments. In this case, even the correct matter density scaling factor
output is used.

The example on page~\pageref{ex:corrproj} demonstrates how one can obtain \figu{projex} (right) with keeping all parameters but $\deltacp$ fixed, as well as how one can include the full $n$-parameter correlation with external input. It also demonstrates how these two compare to each other. One can easily read off this example that 
there is a substantial impact of the
correlation with oscillation parameters other than $\deltacp$. 
Note that it uses the function
\GLB{glbChiNP} for arbitrary projections from the next section for the minimization over $\deltacp$.
In addition, there is one interesting feature in guessing the
oscillation parameters in this example: In order to avoid falling
into the wrong minimum, the fit value of $\deltacp$ is guessed from \figu{projex} (left). This quite sophisticated ``guessing'' is typical
for neutrino factories because of the $(\deltacp, \theta_{13})$-degeneracy, whereas it is for superbeams often sufficient
to use the best-fit values. A strong indication for a wrong guessing 
are discontiuous jumps in the projected $\chi^2$-function, where the minimizer jumps from one minimum to another. In such cases, the starting point of the minimizer has to be adjusted to help it find the true minimum.

Other examples for projections onto a parameter axis while keeping exactly one parameter fixed are \GLB{glbChiTheta23}, \GLB{glbChiDm}, and \GLB{glbChiDms}, which
can be found in \Tab~\ref{tab:stdfunctions} on page~\pageref{tab:stdfunctions}.
